---
title: "p8105_hw1_mk4022.Rmd"
author: "Mary Knoop"
date: "9/17/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

HW 1 Problem 1:
Part A
Create a data frame comprised of:

a random sample of size 8 from a standard Normal distribution,
a logical vector indicating whether elements of the sample are greater than 0,
a character vector of length 8, and
a factor vector of length 8, with 3 different factor “levels”


```{r create_df}
#create dataframe
library(tidyverse)
set.seed(1234)
df_problem1 = tibble(
  norm_samp = rnorm(8),
  log_vec = norm_samp > 0,
  char_vec = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "January", "February", "March"),
  factor_vec = factor(c("day","week","month","day","week","month","day","week"))
)
```

HW 1 Problem 1:
Part B

Try to take the mean of each variable in your dataframe

Note we cannot obtain the mean for char_vec and factor_vec since they are not numeric variables

```{r means}
# take means of each variable
mean(pull(df_problem1, norm_samp))
mean(pull(df_problem1, log_vec))
mean(pull(df_problem1, char_vec))
mean(pull(df_problem1, factor_vec))
```

HW1 Problem 1: 
Part C
Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output).

Note that converting the log_vec produces an output with all 0's and 1's, converting char_vec produces an error, and converting factor_vec transforms the various inputs into 1's, 2's, and 3's corresponding to their matching factor. 

```{r eval = FALSE}
#convert variables without showing output using eval = False to begin new chunk
as.numeric(pull(df_problem1,log_vec))
as.numeric(pull(df_problem1,char_vec))
as.numeric(pull(df_problem1,factor_vec))
```


HW1 Problem 1: 
Part D
In a second code chunk:

convert the logical vector to numeric, and multiply the random sample by the result,
convert the logical vector to a factor, and multiply the random sample by the result,
convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

Note that converting the log_vec to a factor and multiplying produces an error 'not meaningful for factors' since factors are not numeric variables. 

```{r convert_and_multiply}
as.numeric(pull(df_problem1, log_vec))*(pull(df_problem1, norm_samp))
as.factor(pull(df_problem1, log_vec))*(pull(df_problem1, norm_samp))
as.numeric((as.factor(pull(df_problem1, log_vec))))*(pull(df_problem1, norm_samp))
```





